How topos theory can help commutative algebra

### 2018-07-09 slides-como2018.tex
typo

\begin{frame}{The Kripke--Joyal semantics of~$\boldsymbol{\Sh(X)}$}
  \small\vspace*{-0.5em}
  Let~$X$ be a topological space. We recursively define
  \[ U \models \varphi \quad \text{(``$\varphi$ holds on~$U$'')} \]
  \mbox{for open subsets~$U \subseteq X$ and formulas~$\varphi$.
  Write~``$\Sh(X) \models \varphi$'' to mean~$X \models \varphi$.}
  \footnotesize
  \[ \renewcommand{\arraystretch}{1.08}\begin{array}{@{}l@{\ }c@{\ }l@{}}
  U \models \top &\Ll& \text{true} \\
  U \models \bot &\Ll& \hcancel{\text{false}}{0pt}{3pt}{0pt}{-2pt}\ U = \emptyset \\
  U \models s = t \? F &\Ll& s|_U = t|_U \in F(U) \\
  U \models \varphi \wedge \psi &\Ll&
  \text{$U \models \varphi$ and $U \models \psi$} \\
  U \models \varphi \vee \psi &\Ll&
  \hcancel{\text{$U \models \varphi$ or $U \models \psi$}}{0pt}{3pt}{0pt}{-2pt}\ \text{there exists a covering $U = \bigcup_i U_i$} \\
  && \quad\quad\text{such that for all~$i$: $U_i \models \varphi$ or $U_i \models \psi$} \\
  U \models \varphi \Rightarrow \psi &\Ll&
  \text{for all open~$V \subseteq U$: }
  \text{$V \models \varphi$ implies $V \models \psi$} \\
  U \models \forall s \? F\_ \varphi(s) &\Ll&
  \text{for all open $V \subseteq U$ and sections~$s_0 \in F(V)$: $V \models \varphi(s_0)$} \\
  U \models \forall F\_ \varphi(F) &\Ll&
  \text{for all open $V \subseteq U$ and sheaves~$F_0$ over~$V$: $V \models \varphi(F_0)$} \\
  U \models \exists s \? F\_ \varphi(s) &\Ll&
  \hcancel{\text{there exists $s_0 \in F(U)$ such that $U \models \varphi(s_0)$}}{0pt}{3pt}{0pt}{-2pt} \\
  && \quad\quad
  \text{there exists a covering $U = \bigcup_i U_i$ such that for all~$i$:} \\
  && \quad\quad\quad\quad \text{there exists~$s_0 \in F(U_i)$ such that $U_i \models \varphi(s_0)$} \\
  U \models \exists F\_ \varphi(F) &\Ll&
  \hcancel{\text{there exists a sheaf $F_0$ on $U$ such that $U \models
  \varphi(F_0)$}}{0pt}{3pt}{0pt}{-2pt} \\
  && \quad\quad
  \text{there exists a covering $U = \bigcup_i U_i$ such that for all~$i$:} \\
  && \quad\quad\quad\quad \text{there exists a sheaf~$F_0$ on $U_i$ such that $U_i \models \varphi(F_0)$}
  \end{array} \]

@@

  \jnote{1}{Toposes, and also more specifically classifying toposes, originated
  in algebraic geometry. It is therefore deeply embarrassing that as of now,
  still very little is known about the theories classified by the major toposes
  in active use by algebraic geometers.
  \begin{center}
    \includegraphics[width=0.99\textwidth]{wraith-generic-galois-theory} \\
    \footnotesize
    Gavin Wraith. Generic Galois theory of local rings. 1979.
  \end{center}}

>>>>

\begin{frame}{The Kripke--Joyal semantics of~$\boldsymbol{\Sh(X)}$}
  \small\vspace*{-0.5em}
  Let~$X$ be a topological space. We recursively define
  \[ U \models \varphi \quad \text{(``$\varphi$ holds on~$U$'')} \]
  \mbox{for open subsets~$U \subseteq X$ and formulas~$\varphi$.
  Write~``$\Sh(X) \models \varphi$'' to mean~$X \models \varphi$.}
  \footnotesize
  \[ \renewcommand{\arraystretch}{1.08}\begin{array}{@{}l@{\ \ }c@{\ \ }l@{}}
  U \models \top &\Ll& \text{true} \\
  U \models \bot &\Ll& \hcancel{\text{false}}{0pt}{3pt}{0pt}{-2pt}\ U = \emptyset \\
  U \models s = t \? F &\Ll& s|_U = t|_U \in F(U) \\
  U \models \varphi \wedge \psi &\Ll&
  \text{$U \models \varphi$ and $U \models \psi$} \\
  U \models \varphi \vee \psi &\Ll&
  \hcancel{\text{$U \models \varphi$ or $U \models \psi$}}{0pt}{3pt}{0pt}{-2pt}\ \text{there exists a covering $U = \bigcup_i U_i$} \\
  && \quad\quad\text{such that for all~$i$: $U_i \models \varphi$ or $U_i \models \psi$} \\
  U \models \varphi \Rightarrow \psi &\Ll&
  \text{for all open~$V \subseteq U$: }
  \text{$V \models \varphi$ implies $V \models \psi$} \\
  U \models \forall s \? F\_ \varphi(s) &\Ll&
  \text{for all open $V \subseteq U$ and sections~$s_0 \in F(V)$: $V \models \varphi(s_0)$} \\
  U \models \forall F\_ \varphi(F) &\Ll&
  \text{for all open $V \subseteq U$ and sheaves~$F_0$ over~$V$: $V \models \varphi(F_0)$} \\
  U \models \exists s \? F\_ \varphi(s) &\Ll&
  \hcancel{\text{there exists $s_0 \in F(U)$ such that $U \models \varphi(s_0)$}}{0pt}{3pt}{0pt}{-2pt} \\
  &&
  \text{there exists a covering $U = \bigcup_i U_i$ such that for all~$i$:} \\
  && \quad\quad \text{there exists~$s_0 \in F(U_i)$ such that $U_i \models \varphi(s_0)$} \\
  U \models \exists F\_ \varphi(F) &\Ll&
  \hcancel{\text{there exists a sheaf $F_0$ on $U$ such that $U \models
  \varphi(F_0)$}}{0pt}{3pt}{0pt}{-2pt} \\
  &&
  \text{there exists a covering $U = \bigcup_i U_i$ such that for all~$i$:} \\
  && \quad\quad \text{there exists a sheaf~$F_0$ on $U_i$ such that $U_i \models \varphi(F_0)$}
  \end{array} \]

@@

  \jnote{1}{Toposes, and also more specifically classifying toposes, originated
  in algebraic geometry. It is therefore deeply embarrassing that as of now,
  still very little is known about the theories classified by the major toposes
  in active use by algebraic geometers.
  \begin{center}
    \includegraphics[width=0.99\textwidth]{wraith-generic-galois-theory} \\
    \footnotesize
    Gavin Wraith. Generic Galois theory of local rings. 1979. \\
    (A cult classic and must-read for anyone interested in \\
    the intersection of topos theory and algebraic geometry.)
  \end{center}}
